# Census and ACS Data Analysis
Author: Phuong Tseng
<p>Date: "April 3, 2018"</p>

**This script analyzes various census and acs indicators to understand the distribution of opportunity in the Bay Area. These ACS indicators are adult educational attainment, public assistance, commute time, employment rate, poverty rate, and health insurance. The two variables that require decennial census data are poverty and racial segregation indicators that we use as filters.**

#Load all files into myfiles
myfiles = function(mypath){
  filenames=list.files(path=mypath, full.names=TRUE)
  datalist = lapply(filenames, function(x){
    read.csv(file=x,header=T)
  })
  Reduce(function(x,y) {merge(x,y)}, datalist)}

#do this merge of all my files
mymergedfiles = myfiles("/Users/ptseng/Documents/Research/Historical Race Data/native/mergeme")



### 1. Set-up the libraries and directory
```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir=normalizePath("/Users/ptseng/Desktop/OM/Testing_R packages/ACS_16_5YR"))
knitr::opts_chunk$set(echo = F, , warning=FALSE)
x <- c("sp", "dplyr", "tidyr", "plyr", "data.table", "readr", "bit", "bit64", "stringr", "tigris", "acs", "stringr", "tmap", "tidycensus", "readxl")
lapply(x, require, character.only=TRUE)
```

#### Handling Spatial Data using Tigris
```{r Handling Spatial Data, eval=FALSE, include=FALSE}
bayarea_counties <- c(1,13,41,55,75,81,85,95,97)
tracts <- tracts(state = 'CA', county = c(1,13,41,55,75,81,85,95,97), cb=TRUE)
#api.key.install(key="YOUR KEY")
bayarea<-acs::geo.make(state=c("CA"), county=c(1,13,41,55,75,81,85,95,97), tract="*")
```
The counties column above corresponds with Bay Area 9 counties:
1 - Alameda, 13 - Contra Costa, 41 - Marin, 55 - Napa, 75 - SF, 81 - San Mateo, 85 - Santa Clara, 95 - Solano, 97 - Sonoma

#### ACS Package
```{r ACS Package, eval=FALSE, include=FALSE}
adedu_ca<-acs.fetch(endyear = 2012, span = 5, geography = bayarea, table.number="S1501", col.names = "pretty", case.sensitive=F)
```

Note: The ACS package does not get data beyond 2012? Not updated?
Use of col.names = "pretty" above gives the full column definitions, so if you want Census variable IDs use col.names="auto".


### 2. Read the data
```{r Read Data, echo=FALSE}
# Education Domain: Adult Educational Attainment
adulted <- fread("data/input/ACS_16_5YR_S1501_with_ann.csv")

# Economic & Mobility: Public Assistance 
passist <- fread("data/input/ACS_16_5YR_B19057_with_ann.csv")

# Economic & Mobility: Median Home Value
mhval <- fread("data/input/ACS_16_5YR_B25077_with_ann.csv")

# Economic & Mobility: Commute Time
#comtime <- fread("data/input/ACS_16_5YR_B08303_with_ann.csv")

# Economic & Mobility: Employment
employed <- fread("data/input/ACS_16_5YR_B23001_with_ann.csv")

#Housing & Neighborhood Domain: Neighborhood Poverty Rate 200FPL
povr <- fread("data/input/ACS_16_5YR_C17002_with_ann.csv")

#Housing & Neighborhood Domain: Health Insurance
insurance <- fread("data/input/ACS_16_5YR_S2701_with_ann.csv")

owneroccupied <- fread("data/input/ACS_16_5YR_B25008_with_ann.csv")

#Decennial Census Race/Ethnicity table
census_race <- fread("data/input/DEC_10_SF1_SF1DP1.csv", skip = 2, 
  header = FALSE,
  col.names =colnames(fread("data/input/DEC_10_SF1_SF1DP1.csv", nrows=0)))

#ACS Race table
acs_race <- fread("data/input/ACS_16_5YR_B03002.csv", skip = 2, 
  header = FALSE,
  col.names =colnames(fread("data/input/ACS_16_5YR_B03002.csv", nrows=0)))


```

### Decennial Census
```{r Decennial Census Race/Ethnicity, echo=FALSE}
census_race <- select(census_race, longID=GEO.id, tractID=GEO.id2,
  total_pop_2010=HD01_S106, hispanic_2010=HD01_S107,
  pct_hispanic_2010=HD02_S107, white_2010=HD01_S123,
  pct_white_2010=HD02_S123, black_2010=HD01_S124,
  pct_black_2010=HD02_S124,
  asian_2010=HD01_S126, pct_asian_2010=HD02_S126,
  amer_indian_2010=HD01_S125, hawaiian_2010=HD01_S127,
  other_race_2010=HD01_S128, two_races_2010=HD01_S129) %>% 
  mutate(tot_other_race_2010 = (amer_indian_2010 + hawaiian_2010 +
      other_race_2010 + two_races_2010), pct_tot_other_race_2010 =
      (tot_other_race_2010/total_pop_2010))

acs_race <-  select(acs_race, longID=GEO.id, tractID=GEO.id2,
  total_pop_2016=HD01_VD01, moe_total_pop_2016=HD02_VD01,
  white_2016=HD01_VD03, moe_white_2016=HD02_VD03, black_2016=HD01_VD04,
  moe_black_2016=HD02_VD04, amer_indian_2016=HD01_VD05,
  moe_amer_indian_2016=HD02_VD05, asian_2016=HD01_VD06,
  moe_asian_2016=HD02_VD06, hawaiian_2016=HD01_VD07,
  moe_hawaiian_2016=HD02_VD07, other_race_2016=HD01_VD08,
  moe_other_race_2016=HD02_VD08, two_races_2016=HD01_VD09,
  moe_two_races_2016=HD02_VD09, hispanic_2016=HD01_VD12,
  moe_hispanic_2016=HD02_VD12) %>%
  # variable and MOE for total of "other" races
  mutate(tot_other_race_2016 = (amer_indian_2016 + hawaiian_2016 + other_race_2016 + two_races_2016))

# loop through rows to create MOE for total of other races
for(i in 1:nrow(acs_race)){
  row = acs_race[i,]
acs_race$moe_tot_other_race_2016[i] <- moe_sum(
  c(row$moe_amer_indian_2016,row$moe_hawaiian_2016,
    row$moe_other_race_2016, row$moe_two_races_2016), 
  c(row$amer_indian_2016, row$hawaiian_2016,
    row$other_race_2016, row$two_races_2016), 
  na.rm = TRUE) 
}

#select only relevant variables
census_race <- select(census_race, -amer_indian_2010:-two_races_2010)
acs_race <- select(acs_race, -hawaiian_2016:-moe_two_races_2016)
```

### Education Domain
#### i. Adult Educational Attainment
```{r Adult Educational Attainment S1501, echo=FALSE}
adulted <- select(adulted, GEO.id, GEO.id2, HC01_EST_VC14, HC01_EST_VC15, HC01_MOE_VC14, HC01_MOE_VC15, HC02_EST_VC18, HC02_MOE_VC18)

colnames(adulted) <- c("longID", "tractID", "est_ba_2016", "est_gradprof_2016", "moe_ba_2016", "moe_gradprof_2016",   "p_baplus_2016", "moe_p_baplus_2016")

adulted <- adulted[-c(1),,drop=F]

adulted_2016 <- adulted
adulted_2016 <- mutate(adulted_2016, longID=as.character(longID), 
  tractID=as.character(tractID), 
  est_ba_2016 = as.numeric(est_ba_2016),
  est_gradprof_2016 = as.numeric(est_gradprof_2016),
  moe_ba_2016 = as.numeric(moe_ba_2016),
  moe_gradprof_2016 = as.numeric(moe_gradprof_2016),
  p_baplus_2016 = as.numeric(p_baplus_2016),
  moe_p_baplus_2016 = as.numeric(moe_p_baplus_2016),
  # Convert percentage of baplus to coefficient of variation for reliability scores
  cv_p_baplus_2016 = (moe_p_baplus_2016/1.645)/p_baplus_2016,
  Reliability = if_else(cv_p_baplus_2016 > 0.3, 'Unreliable', 'Reliable',
    missing = 'NA')
  ) 
glimpse(adulted_2016)
nrow(adulted_2016)

# Extract Bay Area Counties
adulted_2016$CountyID <- substr(adulted_2016$tractID, 1,5)

bayarea_adulted_2016 <- filter(adulted_2016, CountyID=="06001" | CountyID=="06013" | CountyID=="06041" | CountyID=="06055" | CountyID=="06075" | CountyID=="06081" | CountyID=="06085" | CountyID=="06095" | CountyID=="06097")

nrow(bayarea_adulted_2016)

# Descriptive Statistics
#stats_adulted_2016 <- summarise(adulted_2016, mean_p_baplus_2016 = mean(adulted_2016$p_baplus_2016, na.rm=T), sd_p_baplus_2016=sd(adulted_2016$p_baplus_2016, na.rm = T))

bayarea_adulted_2016$est_baplus_2016 <- bayarea_adulted_2016$est_ba_2016 + bayarea_adulted_2016$est_gradprof_2016

bayarea_stats_adulted_2016 <- summarise(bayarea_adulted_2016, mean_p_baplus_2016 = mean(p_baplus_2016, na.rm=T), sd_p_baplus_2016=sd(p_baplus_2016, na.rm = T))

bayarea_adulted_2016$zscores <- (bayarea_adulted_2016$p_baplus_2016 - bayarea_stats_adulted_2016$mean_p_baplus_2016)/bayarea_stats_adulted_2016$sd_p_baplus_2016


```

### Economic & Mobility Domain
#### i. Public Assistance
```{r Public Assistance B19057, echo=FALSE}
passist <- select(passist, GEO.id, GEO.id2, HD01_VD01, HD01_VD02, HD02_VD01, HD02_VD02)

colnames(passist) <- c("longID", "tractID", "est_total_2016", "est_passist_2016", "moe_total_2016", "moe_passist_2016")

passist <- passist[-c(1),,drop=F]

passist_2016 <- passist
passist_2016 <- mutate(passist_2016, longID=as.character(longID), 
  tractID=as.character(tractID), 
  est_total_2016 = as.numeric(est_total_2016),
  est_passist_2016 = as.numeric(est_passist_2016),
  moe_total_2016 = as.numeric(moe_total_2016),
  moe_passist_2016 = as.numeric(moe_passist_2016),
  # Convert MOE to coefficient of variation for reliability scores
  cv_est_passist_2016 = (moe_passist_2016/1.645)/est_passist_2016,
  Reliability = if_else(cv_est_passist_2016 > 0.3, 'Unreliable',
    'Reliable', missing = 'NA')
  ) 
glimpse(passist_2016)
nrow(passist_2016)

# Extract Bay Area Counties
passist_2016$CountyID <- substr(passist_2016$tractID, 1,5)

bayarea_passist_2016 <- filter(passist_2016, CountyID=="06001" | CountyID=="06013" | CountyID=="06041" | CountyID=="06055" | CountyID=="06075" | CountyID=="06081" | CountyID=="06085" | CountyID=="06095" | CountyID=="06097")

nrow(bayarea_passist_2016)

#Descriptive Statistics
#Percentage of people with public assistance
bayarea_passist_2016$EC_PASSIST <- bayarea_passist_2016$est_passist_2016/bayarea_passist_2016$est_total_2016

#Descriptive Statistics of people with public assistance in the bay area
bayarea_stats_passist_2016 <- summarise(bayarea_passist_2016, mean_EC_PASSIST = mean(bayarea_passist_2016$EC_PASSIST, na.rm=T), sd_EC_PASSIST=sd(bayarea_passist_2016$EC_PASSIST, na.rm = T))

bayarea_passist_2016$zscores <- (bayarea_passist_2016$EC_PASSIST - bayarea_stats_passist_2016$mean_EC_PASSIST)/bayarea_stats_passist_2016$sd_EC_PASSIST

bayarea_passist_2016$adj_zscores <- bayarea_passist_2016$zscores * -1

#summary(bayarea_passist_2016) #To check number of NA's


```

#### ii. Median Home Value
```{r Median Home Value B25077, echo=FALSE}
# Clean the B25077 data
mhval <- select(mhval, GEO.id, GEO.id2, HD01_VD01, HD02_VD01)

colnames(mhval) <- c("longID", "tractID", "est_mhval_2016", "moe_mhval_2016")

mhval <- mhval[-c(1),,drop=F]

mhval_2016 <- mhval
mhval_2016 <- mutate(mhval_2016, longID=as.character(longID), 
  tractID=as.character(tractID), 
  est_mhval_2016 = as.numeric(est_mhval_2016),
  moe_mhval_2016 = as.numeric(moe_mhval_2016),
  # Convert MOE to coefficient of variation for reliability scores
  cv_est_mhval_2016 = (moe_mhval_2016/1.645)/est_mhval_2016,
  Reliability = if_else(cv_est_mhval_2016 > 0.3, 'Unreliable',
    'Reliable', missing = 'NA')
  ) 
               
glimpse(mhval_2016)
nrow(mhval_2016)

# Extract
mhval_2016$CountyID <- substr(mhval_2016$tractID, 1,5)

bayarea_mhval_2016 <- filter(mhval_2016, CountyID=="06001" | CountyID=="06013" | CountyID=="06041" | CountyID=="06055" | CountyID=="06075" | CountyID=="06081" | CountyID=="06085" | CountyID=="06095" | CountyID=="06097")

nrow(bayarea_mhval_2016)

#Descriptive Statistics
# Median Home Value in the bay area
bayarea_stats_mhval_2016 <- summarise(bayarea_mhval_2016, mean_est_mhval_2016 = mean(est_mhval_2016, na.rm=T), sd_est_mhval_2016=sd(est_mhval_2016, na.rm = T))

bayarea_mhval_2016$zscores <- (bayarea_mhval_2016$est_mhval_2016 - bayarea_stats_mhval_2016$mean_est_mhval_2016)/bayarea_stats_mhval_2016$sd_est_mhval_2016


#summary(bayarea_mhval_2016) #To check number of NA's
```

#### iii. Commute Time, Needs column names
```{r Commute Time B08303, eval=FALSE, include=FALSE}

comtime <- select(comtime, GEO.id, GEO.id2 )

colnames(comtime) <- c("longID", "tractID", "EST_TOT",
"MOE_TOT",
"EST_30_34",
"MOE_30_34",
"EST_35_39",
"MOE_35_39",
"EST_40_44",
"MOE_40_44",
"EST_45_59",
"MOE_45_59",
"EST_60_89",
"MOE_60_89",
"EST_90_MOR",
"MOE_90_MOR",
"AGG_30_MOR",
"MOE_30_MOR",
"EC_COMTIM",
"MOE_30_TOT")

comtime <- comtime[-c(1),,drop=F]

comtime_2016 <- comtime

#Extract 
#comtime_2016$CountyID <- substr(comtime_2016$tractID, 1,5)

#bayarea_comtime_2016 <- filter(comtime_2016, CountyID=="06001" | CountyID=="06013" | CountyID=="06041" | CountyID=="06055" | CountyID=="06075" | CountyID=="06081" | CountyID=="06085" | CountyID=="06095" | CountyID=="06097")

#nrow(bayarea_comtime_2016)

#mean_and_sd <- summarise(comtime_2016, mean_p_baplus_2016 = mean(comtime_2016$p_baplus_2016, na.rm=T), sd_p_baplus_2016=sd(comtime_2016$p_baplus_2016, na.rm = T))

#ba_mean_and_sd <- summarise(bayarea_comtime_2016, mean_p_baplus_2016 = mean(bayarea_comtime_2016$p_baplus_2016, na.rm=T), sd_p_baplus_2016=sd(bayarea_comtime_2016$p_baplus_2016, na.rm = T))

#bayarea_comtime_2016$zscores <- (bayarea_comtime_2016$p_baplus_2016 - ba_mean_and_sd$mean_p_baplus_2016)/ba_mean_and_sd$sd_p_baplus_2016

#bayarea_comtime_2016$est_baplus_2016 <- bayarea_comtime_2016$est_ba_2016 + bayarea_comtime_2016$est_gradprof_2016

#summary(bayarea_comtime_2016) #To check number of NA's
```

#### iv. Employment
```{r Employment B23001, echo=FALSE}
employed <- select(employed, GEO.id, GEO.id2, 
HD01_VD10, #20-21 Male total
HD02_VD10, 
HD01_VD12, #20-21 Male Armed forces
HD02_VD12, 
HD01_VD14, #20-21 Male Civilian Employed
HD02_VD14, 
HD01_VD17, #22-24 Male Total
HD02_VD17,
HD01_VD19, #22-24 Male Armed forces
HD02_VD19,
HD01_VD21, #22-24 Male civilian employed
HD02_VD21,
HD01_VD24, #25-29 Male total
HD02_VD24,
HD01_VD26, #25-29 Male armed forces
HD02_VD26,
HD01_VD28, #25-29 Male civilian employed
HD02_VD28,
HD01_VD31, #30-34 Male total
HD02_VD31,
HD01_VD33, #30-34 Male armed forces
HD02_VD33,
HD01_VD35 , #30-34 Male civilian employed
HD02_VD35 ,
HD01_VD38 , #35-44 Male total
HD02_VD38 ,
HD01_VD40 , #35-44 Male armed forces
HD02_VD40 ,
HD01_VD42 , #35-44 Male civilian employed
HD02_VD42 ,
HD01_VD45 , #45-54 Male total
HD02_VD45 ,
HD01_VD47 , #45-54 Male armed forces
HD02_VD47 ,
HD01_VD49 , #45-54 Male civilian employed
HD02_VD49 ,
HD01_VD52 , #55-59 Male total
HD02_VD52 ,
HD01_VD54 , #55-59 Male armed forces
HD02_VD54 ,
HD01_VD56 , #55-59 Male civilian employed
HD02_VD56 ,
HD01_VD99 , #20-21 Female total
HD02_VD99 ,
HD01_VD101 , #20-21 Female armed forces
HD02_VD101 ,
HD01_VD103 , #20-21 Female civilian employed
HD02_VD103 ,
HD01_VD106 , #22-24 Female total
HD02_VD106 ,
HD01_VD108 , #22-24 Female armed forces
HD02_VD108 ,
HD01_VD110 , #22-24 Female civilian employed
HD02_VD110 ,
HD01_VD113 , #25-29 Female total
HD02_VD113 ,
HD01_VD115 , #25-29 Female armed forces
HD02_VD115 ,
HD01_VD117 , #25-29 Female civilian employed
HD02_VD117 ,
HD01_VD120 , #30-34 Female total
HD02_VD120 ,
HD01_VD122 , #30-34 Female armed forces
HD02_VD122 ,
HD01_VD124 , #30-34 Female civilian employed
HD02_VD124 ,
HD01_VD127 , #35-44 Female total
HD02_VD127 ,
HD01_VD129 , #35-44 Female armed forces
HD02_VD129 ,
HD01_VD131 , #35-44 Female civilian employed
HD02_VD131 ,
HD01_VD134 , #45-54 Female total
HD02_VD134 ,
HD01_VD136 , #45-54 Female armed forces
HD02_VD136 ,
HD01_VD138 , #45-54 Female civilian employed
HD02_VD138 ,
HD01_VD141 , #55-59 Female total
HD02_VD141 ,
HD01_VD143 , #55-59 Female armed forces
HD02_VD143 ,
HD01_VD145 , #55-59 Female civilian employed 
HD02_VD145)

#convert all HD* variables to numeric
employed <- mutate_at(employed, vars(starts_with('HD')), as.numeric)

#For this specific indicator, let's take the bay area estimates instead of the entire california
# Extract for bay area
employed_2016 <- employed
employed_2016$CountyID <- substr(employed_2016$GEO.id2, 1,5)
employed_2016 <- employed_2016[-c(1),,drop=F]

bayarea_employed_2016 <- filter(employed_2016, CountyID=="06001" | CountyID=="06013" | CountyID=="06041" | CountyID=="06055" | CountyID=="06075" | CountyID=="06081" | CountyID=="06085" | CountyID=="06095" | CountyID=="06097")

nrow(bayarea_employed_2016)

# Population Ages 20-60
bayarea_employed_2016 <- mutate(bayarea_employed_2016, pop_20to60_2016 = HD01_VD10 + HD01_VD17 + HD01_VD24 + HD01_VD31 + HD01_VD38 + HD01_VD45 + HD01_VD52 + HD01_VD99 + HD01_VD106 + HD01_VD113 + HD01_VD120 + HD01_VD127 + HD01_VD134 + HD01_VD141)

# bayarea_employed_2016 Civilian and Armed Forces, 20 to 60
bayarea_employed_2016 <- mutate(bayarea_employed_2016, employed_pop_20to60_2016 = HD01_VD12 + HD01_VD14 +  HD01_VD19 + HD01_VD21 +  HD01_VD26 + HD01_VD28 + HD01_VD33 + HD01_VD35 + HD01_VD40 + HD01_VD42 + HD01_VD47 + HD01_VD49 + HD01_VD54 + HD01_VD56 + HD01_VD101 + HD01_VD103 + HD01_VD108 + HD01_VD110 + HD01_VD115 + HD01_VD117 + HD01_VD122 + HD01_VD124 + HD01_VD129 + HD01_VD131 + HD01_VD136 + HD01_VD138 + HD01_VD143 + HD01_VD145)

# MOE Population Ages 20-60
for(i in 1:nrow(bayarea_employed_2016)){
  # MOE Population Ages 20-60
  row <- bayarea_employed_2016[i,]
bayarea_employed_2016$moe_pop_20to60_2016[i] <- moe_sum(
  c(row$HD02_VD10, row$HD02_VD17, row$HD02_VD24, row$HD02_VD31,
    row$HD02_VD38, row$HD02_VD45, row$HD02_VD52, row$HD02_VD99,
    row$HD02_VD106, row$HD02_VD113, row$HD02_VD120, row$HD02_VD127,
    row$HD02_VD134, row$HD02_VD141),
  c(row$HD01_VD10, row$HD01_VD17, row$HD01_VD24, row$HD01_VD31,
    row$HD01_VD38, row$HD01_VD45, row$HD01_VD52, row$HD01_VD99,
    row$HD01_VD106, row$HD01_VD113, row$HD01_VD120, row$HD01_VD127,
    row$HD01_VD134, row$HD01_VD141),
na.rm = TRUE)

#MOE bayarea_employed_2016 Civilian and Armed Forces, 20-60
bayarea_employed_2016$moe_employed_pop_20to60_2016[i] <- moe_sum(
  c( row$HD02_VD12, row$HD02_VD14,  row$HD02_VD19, row$HD02_VD21,  row$HD02_VD26, row$HD02_VD28, row$HD02_VD33, row$HD02_VD35, row$HD02_VD40, row$HD02_VD42, row$HD02_VD47, row$HD02_VD49, row$HD02_VD54, row$HD02_VD56, row$HD02_VD101, row$HD02_VD103, row$HD02_VD108, row$HD02_VD110, row$HD02_VD115, row$HD02_VD117, row$HD02_VD122, row$HD02_VD124, row$HD02_VD129, row$HD02_VD131, row$HD02_VD136, row$HD02_VD138, row$HD02_VD143, row$HD02_VD145),
  c(row$HD01_VD12, row$HD01_VD14,  row$HD01_VD19, row$HD01_VD21,  row$HD01_VD26, row$HD01_VD28, row$HD01_VD33, row$HD01_VD35, row$HD01_VD40, row$HD01_VD42, row$HD01_VD47, row$HD01_VD49, row$HD01_VD54, row$HD01_VD56, row$HD01_VD101, row$HD01_VD103, row$HD01_VD108, row$HD01_VD110, row$HD01_VD115, row$HD01_VD117, row$HD01_VD122, row$HD01_VD124, row$HD01_VD129, row$HD01_VD131, row$HD01_VD136, row$HD01_VD138, row$HD01_VD143, row$HD01_VD145),
na.rm = TRUE)
}

#"Percent employed in Civilian or Armed Forces, 20-60", and CV for the number of employed 
bayarea_employed_2016 <- mutate(bayarea_employed_2016, pct_employed_20to60_2016 = employed_pop_20to60_2016/pop_20to60_2016,
  cv_employed_2016 =
    (moe_employed_pop_20to60_2016/1.645)/employed_pop_20to60_2016,
  Reliability = if_else(cv_employed_2016 > 0.3, 'Unreliable',
    'Reliable', missing = 'NA')
  )


#"MOE Percent bayarea_employed_2016 in Civilian or Armed Forces, 20-60"
#moe_pct_bayarea_employed_2016_20to60_2016 <-  sqrt(moe_bayarea_employed_2016_pop_20to60_2016^2-(pct_bayarea_employed_2016_20to60_2016^2*moe_pop_20to60_2016^2))/pop_20to60_2016

#Descriptive Statistics
bayarea_stats_employed_2016 <- summarise(bayarea_employed_2016, mean_pct_employed_20to60_2016 = mean(pct_employed_20to60_2016, na.rm=T), sd_pct_employed_20to60_2016=sd(pct_employed_20to60_2016, na.rm = T))

bayarea_employed_2016$zscores <- (bayarea_employed_2016$pct_employed_20to60_2016 - bayarea_stats_employed_2016$mean_pct_employed_20to60_2016)/bayarea_stats_employed_2016$sd_pct_employed_20to60_2016

#summary(bayarea_employed_2016) #To check number of NA's
```

### Housing & Neighborhood Domain

#### i. Neighborhood Poverty Rate 200FPL
```{r Neighborhood Poverty Rate 200FPL, echo=FALSE}
povr <- select(povr, GEO.id, GEO.id2, HD01_VD01, HD02_VD01, HD01_VD08, HD02_VD08)

colnames(povr) <- c("longID", "tractID", "est_total_2016", "moe_total_2016","est_2xFPL_2016","moe_2xFPL_2016")

povr <- povr[-c(1),,drop=F]

povr_2016 <- povr
povr_2016 <- mutate(povr_2016, longID=as.character(longID), 
  tractID=as.character(tractID), 
  est_total_2016 = as.numeric(est_total_2016),
  moe_total_2016 = as.numeric(moe_total_2016),
  est_2xFPL_2016 = as.numeric(est_2xFPL_2016),
  moe_2xFPL_2016 = as.numeric(moe_2xFPL_2016),
  # Convert MOE to coefficient of variation for reliability scores
  cv_2xFPL_2016 =
    (moe_2xFPL_2016/1.645)/est_2xFPL_2016,
  Reliability = if_else(cv_2xFPL_2016 > 0.3, 'Unreliable',
    'Reliable', missing = 'NA'))
               
glimpse(povr_2016)
nrow(povr_2016)

#Extract
povr_2016$CountyID <- substr(povr_2016$tractID, 1,5)

bayarea_povr_2016 <- filter(povr_2016, CountyID=="06001" | CountyID=="06013" | CountyID=="06041" | CountyID=="06055" | CountyID=="06075" | CountyID=="06081" | CountyID=="06085" | CountyID=="06095" | CountyID=="06097")

nrow(bayarea_povr_2016)

#Descriptive Statistics
#Neighborhood poverty rate in the bay area
bayarea_povr_2016$p_2xFPL_2016 <- (bayarea_povr_2016$est_2xFPL_2016/bayarea_povr_2016$est_total_2016)

bayarea_stats_povr_2016 <- summarise(bayarea_povr_2016, mean_p_2xFPL_2016 = mean(p_2xFPL_2016, na.rm=T), sd_p_2xFPL_2016=sd(p_2xFPL_2016, na.rm = T))

bayarea_povr_2016$zscores <- (bayarea_povr_2016$p_2xFPL_2016 - bayarea_stats_povr_2016$mean_p_2xFPL_2016)/bayarea_stats_povr_2016$sd_p_2xFPL_2016

bayarea_povr_2016$adj_zscores <- bayarea_povr_2016$zscores * -1

#summary(bayarea_povr_2016) #To check number of NA's
```

#### ii. Health Insurance
```{r Health Insurance, echo=FALSE}
insurance <- select(insurance, GEO.id, GEO.id2, HC01_EST_VC20,
HC01_MOE_VC20,
HC02_EST_VC20,
HC02_MOE_VC20,
HC04_EST_VC20,
HC04_MOE_VC20,
HC01_EST_VC21,
HC01_MOE_VC21,
HC02_EST_VC21,
HC02_MOE_VC21,
HC04_EST_VC21,
HC04_MOE_VC21)

colnames(insurance) <- c("longID", "tractID", "EST_M",
"MOE_M",
"EST_M_INS",
"MOE_M_INS",
"EST_M_UINS",
"MOE_M_UINS",
"EST_F",
"MOE_F",
"EST_F_INS",
"MOE_F_INS",
"EST_F_UINS",
"MOE_F_UINS")

insurance <- insurance[-c(1), , drop=F]

insurance_2016 <- insurance

insurance_2016 <- mutate(insurance_2016, longID=as.character(longID), 
  tractsID=as.character(tractID)) %>% 
  mutate_at(vars(contains('_')), as.numeric)

glimpse(insurance_2016)
nrow(insurance_2016)

#Extract
insurance_2016$CountyID <- substr(insurance_2016$tractID, 1,5)

bayarea_insurance_2016 <- filter(insurance_2016, CountyID=="06001" | CountyID=="06013" | CountyID=="06041" | CountyID=="06055" | CountyID=="06075" | CountyID=="06081" | CountyID=="06085" | CountyID=="06095" | CountyID=="06097")

nrow(bayarea_insurance_2016)

# MOE for the insured population
for(i in 1:nrow(bayarea_insurance_2016)){
  row <- bayarea_insurance_2016[i,]
bayarea_insurance_2016$moe_Insured[i] <- moe_sum(
  c(row$MOE_M_INS, row$MOE_F_INS),
  c(row$EST_M_INS, row$EST_F_INS),
na.rm = TRUE)
}
#Descriptive Statistics
bayarea_insurance_2016 <- mutate(bayarea_insurance_2016,
  total_pop = EST_M + EST_F, Insured = EST_M_INS + EST_F_INS,
  p_Insured = Insured / total_pop, 
  # Convert insured MOE to coefficient of variation for reliability scores
  cv_Insured = (moe_Insured/1.645)/Insured,
  Reliability = if_else(cv_Insured > 0.3, 'Unreliable',
    'Reliable', missing = 'NA'))


bayarea_stats_insurance_2016 <- summarise(bayarea_insurance_2016, mean_p_Insured = mean(bayarea_insurance_2016$p_Insured, na.rm=T), sd_p_Insured=sd(bayarea_insurance_2016$p_Insured, na.rm = T))

bayarea_insurance_2016$zscores <- (bayarea_insurance_2016$p_Insured - bayarea_stats_insurance_2016$mean_p_Insured)/bayarea_stats_insurance_2016$sd_p_Insured


#summary(bayarea_insurance_2016) #To check number of NA's
```

#### ii. Home Ownership
```{r House Ownership, echo=FALSE}
owneroccupied <- select(owneroccupied, GEO.id, GEO.id2, HD01_VD01, HD02_VD01, HD01_VD02, HD02_VD02, HD01_VD03, HD02_VD03)

colnames(owneroccupied) <- c("longID","tractID","est_total", "moe_total","est_owner","moe_owner","est_renter","moe_renter")

owneroccupied <- owneroccupied[-c(1), , drop=F]

owneroccupied_2016 <- owneroccupied
owneroccupied_2016 <- mutate(owneroccupied_2016, longID=as.character(longID),
  tractID=as.character(tractID), est_total=as.numeric(est_total),
  moe_total=as.numeric(moe_total),est_owner=as.numeric(est_owner),
  moe_owner=as.numeric(moe_owner), est_renter=as.numeric(est_renter),
  moe_renter=as.numeric(moe_renter))

#owneroccupied %>% group_by(owneroccupied$est_owner) %>% summarize(mean_owner = mean(owneroccupied$est_owner))
```

##3. Extract Bay Area Counties
```{r Bay Area Counties: Method 1, echo=FALSE}
#Tried extracting counties with the data.frame function but did not work
#CountyID2 <- data.frame(adulted2$tractID)
#adulted2$COUNTY <- separate(CountyID2, col=1, into=c("County", "Tract"), sep = 5, extra = "drop", fill ="right")
```

```{r Bay Area Counties: Method 2 Show, echo=TRUE}
ls()
```

## 4. Missing and Unreliable Data
```{r Testing Reliability, eval=FALSE, include=FALSE}
#CV <- sd(bayarea_adulted_2016$est_baplus_2016, na.rm=TRUE)/mean(bayarea_adulted_2016$est_baplus_2016, na.rm=TRUE)*100

#bayarea_adulted_2016$Reliablty <- if_else((bayarea_adulted_2016$est_baplus_2016 * 0.3) < CV, "Unreliable", "Reliable", missing="NA")



```

```{r Testing Reliability 2, eval=FALSE, include=FALSE}
#CV <- c(CV, bayarea_stats_adulted_2016$CV, bayarea_stats_insurance_2016$CV)
```

# Assuming that we have a normal distribution, or just plot it
#normal <- rnorm(1583, mean = 1014.83, sd = 870.6773)
#rm(normal)
plot(adulted$p_baplus_2016)
hist(adulted$p_baplus_2016)
```{r save files}
## 5. Export the results:
# Note: Number of records in ACS: 1,588; Number of tracts in shapefile: 1,582; Out of the 1,588 records, nine records are NULL, while six were not in the shapefile and three have tractIDs:
# 
# 1. 06095980000 - Travis Air Force Base in Fairfield, 
# 2. 06081984300 - an airport in San Francisco, and 
# 3. 06075980401 - two Islands southwest of the Bay Area
write_csv(bayarea_adulted_2016, "data/output/bayarea_adulted_2016.csv",na ="NA") 
write_csv(bayarea_employed_2016, "data/output/bayarea_employed_2016.csv",na ="NA") 
write_csv(bayarea_insurance_2016, "data/output/bayarea_insurance_2016.csv",na ="NA") 
write_csv(bayarea_mhval_2016, "data/output/bayarea_mhval_2016.csv",na ="NA") 
write_csv(bayarea_passist_2016, "data/output/bayarea_passist_2016.csv",na ="NA") 
write_csv(bayarea_povr_2016, "data/output/bayarea_povr_2016.csv",na ="NA") 
write_csv(census_race, "data/output/census_race.csv",na ="NA") 
write_csv(bayarea_povr_2016, "data/output/acs_race.csv",na ="NA") 
```

```{r Check to see if the files exist, eval = FALSE}
file.exists("bayarea_adulted_2016.csv","bayarea_employed_2016.csv",
  "bayarea_insurance_2016.csv","bayarea_mhval_2016.csv","bayarea_passist_2016.csv",
  "bayarea_povr_2016.csv")
```
